

\chapter{EXPERIMENT 2B ITEMS} \label{ap:exp2bitems}
\section{New experimental conditions}

\ea Muhtarlar{\i}n/Muhtar{\i}n \c{c}ocu\u{g}u okula yorgun arg{\i}n geldiler/geldi.\\*
({\it The mukhtars/mukhtar's children came\textsubscript{PL/SG} to school exhausted.})
\ex Y\"{o}neticilerin/Y\"{o}neticinin a\c{s}\c{c}{\i}s{\i} mutfakta s\"{u}rekli z{\i}plad{\i}lar/z{\i}plad{\i}.\\*
({\it The executives'/executive's cook jumped\textsubscript{PL/SG} in the kitchen non-stop.})
\ex \"{O}\u{g}retmenlerin/\"{O}\u{g}retmenin m\"{u}d\"{u}r\"{u} s{\i}n{\i}fta birden bay{\i}ld{\i}lar/bay{\i}ld{\i}.\\*
({\it The teachers'/teacher's principal suddenly fainted\textsubscript{PL/SG} in the classroom.})
\ex Mobilyac{\i}lar{\i}n/Mobilyac{\i}n{\i}n marangozu at\"{o}lyeden h{\i}zla uzakla\c{s}t{\i}lar/uzakla\c{s}t{\i}.\\*
({\it The furniture makers'/furniture maker's carpenter hurried\textsubscript{PL/SG} away from the workshop.})
\ex Mahallelilerin/Mahallelinin emlak\c{c}{\i}s{\i} aniden k\"{u}stah\c{c}a g\"{u}ld\"{u}ler/g\"{u}ld\"{u}.\\*
({\it The locals'/local's real estate agent suddenly laughed\textsubscript{PL/SG} arrogantly.})
\ex H\"{u}k\"{u}metlerin/H\"{u}k\"{u}metin akademisyeni sabaha olana kadar a\u{g}lad{\i}lar/a\u{g}lad{\i}.\\*
({\it The governments'/government's academic cried\textsubscript{PL/SG} until morning.})
\ex Projelerin/Projenin ara\c{s}t{\i}rmac{\i}s{\i} arada s{\i}rada s{\i}k{\i}ld{\i}lar/s{\i}k{\i}ld{\i}.\\*
({\it The projects'/project's researcher got\textsubscript{PL/SG} bored every now and then.})
\ex Doktorlar{\i}n/Doktorun hastas{\i} g\"{u}nden g\"{u}ne durmadan zay{\i}flad{\i}lar/zay{\i}flad{\i}.\\*
({\it Doctors'/doctor's patient got\textsubscript{PL/SG} weaker day by day.})
\ex Y\"{o}netmenlerin/Y\"{o}netmenin oyuncusu onikiden \"{o}nce uyudular/uyudu.\\*
({\it The directors'/director's actor fell\textsubscript{PL/SG} asleep before twelve.})
\ex Aristokratlar{\i}n/Aristokrat{\i}n hizmet\c{c}isi yorgun arg{\i}n yatt{\i}lar/yatt{\i}.\\*
({\it The aristocrats'/aristocrat's servant lay\textsubscript{PL/SG} tiredly.})
\ex Konu\c{s}mac{\i}lar{\i}n/Konu\c{s}mac{\i}n{\i}n sunucusu ola\u{g}an\"{u}st\"{u} bir h{\i}zla konu\c{s}tular/konu\c{s}tu.\\*
({\it The speakers'/speaker's presenter spoke\textsubscript{PL/SG} with extraordinary speed.})
\ex M\"{u}zelerin/M\"{u}zenin ziyaret\c{c}isi a\c{c} susuz dola\c{s}t{\i}lar/dola\c{s}t{\i}.\\*
({\it The museums'/museum's visitor wandered\textsubscript{PL/SG} around hungry and thirsty.})
\ex Politikac{\i}lar{\i}n/Politikac{\i}n{\i}n hocas{\i} adliyeden \c{c}abucak \c{c}{\i}kt{\i}lar/\c{c}{\i}kt{\i}.\\*
({\it The politicians'/politician's teacher quickly left\textsubscript{PL/SG} the courthouse.})
\ex Hakimlerin/Hakimin \c{c}ayc{\i}s{\i} nedensiz yere k{\i}zd{\i}lar/k{\i}zd{\i}.\\*
({\it The judges'/judge's tea man got\textsubscript{PL/SG} angry for no reason.})
\ex Oyuncular{\i}n/Oyuncunun hem\c{s}iresi etrafta ama\c{c}s{\i}zca gezdiler/gezdi.\\*
({\it The players'/player's nurse wandered\textsubscript{PL/SG} aimlessly.})
\ex \c{C}al{\i}\c{s}anlar{\i}n/\c{C}al{\i}\c{s}an{\i}n m\"{u}diresi biraz \"{o}nce arad{\i}lar/arad{\i}.\\*
({\it The workers'/worker's manager just called\textsubscript{PL/SG}.})
\ex Milyonerlerin/Milyonerin terzisi tamamen gereksizce ba\u{g}{\i}rd{\i}lar/ba\u{g}{\i}rd{\i}.\\*
({\it The millionaires'/millionaire's tailor yelled\textsubscript{PL/SG} completely unnecessary.})
\ex Bebeklerin/Bebe\u{g}in bak{\i}c{\i}s{\i} \c{c}ok kibar davrand{\i}lar/davrand{\i}.\\*
({\it The babies'/baby's sitter was/were very kind.})
\ex Kom\c{s}ular{\i}n/Kom\c{s}unun dad{\i}s{\i} sahil boyunca s\"{u}z\"{u}ld\"{u}ler/s\"{u}z\"{u}ld\"{u}.\\*
({\it The neighbors'/neighbor's nanny glided\textsubscript{PL/SG}along the beach.})
\ex Polislerin/Polisin tamircisi a\c{c} susuz \c{c}al{\i}\c{s}t{\i}lar/\c{c}al{\i}\c{s}t{\i}.\\*
({\it The cops'/cop's mechanic worked\textsubscript{PL/SG} hungry and thirsty.})
\ex Modac{\i}lar{\i}n/Modac{\i}n{\i}n taksicisi saatlerce durmaks{\i}z{\i}n i\c{c}tiler/i\c{c}ti.\\*
({\it The fashionistas'/fashionista's taxi driver drank\textsubscript{PL/SG} non-stop for hours.})
\ex Sanat\c{c}{\i}lar{\i}n/Sanat\c{c}{\i}n{\i}n \c{c}alg{\i}c{\i}s{\i} feci bir \c{s}ekilde \"{o}ld\"{u}ler/\"{o}ld\"{u}.\\*
({\it The artists'/artist's instrumentalist died\textsubscript{PL/SG} tragically.})
\ex Dedektiflerin/Dedektifin di\c{s}\c{c}isi ilk kez \c{c}{\i}lg{\i}nca e\u{g}lendiler/e\u{g}lendi.\\*
({\it The detectives'/detective's dentist had\textsubscript{PL/SG} a blast for the first time.})
\ex Esnaflar{\i}n/Esnaf{\i}n m\"{u}\c{s}terisi \c{s}ikayettten hemen sonra sustular/sustu.\\*
({\it The shopkeepers'/shopkeeper's customer fell\textsubscript{PL/SG} silent immediately after complaining.})
\ex \c{S}ark{\i}c{\i}lar{\i}n/\c{S}ark{\i}c{\i}n{\i}n korumas{\i} her zamanki gibi geciktiler/gecikti.\\*
({\it The singers'/singer's bodyguard was/were delayed as usual.})
\ex G\"{o}stericilerin/G\"{o}stericinin izleyicisi ak\c{s}ama kadar sessizce oturdular/oturdu.\\*
({\it The demonstrators'/demonstrator's audience sat\textsubscript{PL/SG} in silence until the evening.})
\ex Cerrahlar{\i}n/Cerrah{\i}n hastas{\i} ak\c{s}amki g\"{o}steriden \"{o}nce ka\c{c}t{\i}lar/ka\c{c}t{\i}.\\*
({\it The surgeons'/surgeon's patient fled\textsubscript{PL/SG} before the evening's performance.})
\ex \"{O}\u{g}rencilerin/\"{O}\u{g}rencinin terapisti bile bile ge\c{c} kald{\i}lar/kald{\i}.\\*
({\it The students'/student's therapist was/were deliberately late.})
\ex Fabrikat\"{o}rlerin/Fabrikat\"{o}r\"{u}n i\c{s}\c{c}isi beklenmedik bir anda hastaland{\i}lar/hastaland{\i}.\\*
({\it The fabricators'/fabricator's worker fell\textsubscript{PL/SG} ill unexpectedly.})
\ex Yolcular{\i}n/Yolcunun hostesi poyrazdan dolay{\i} \"{u}\c{s}\"{u}d\"{u}ler/\"{u}\c{s}\"{u}d\"{u}.\\*
({\it The passengers'/passenger's stewardess was/were cold because of the north wind.})
\ex \c{S}of\"{o}rlerin/\c{S}of\"{o}r\"{u}n yolcusu yemekten sonra yine ac{\i}kt{\i}lar/ac{\i}kt{\i}.\\*
({\it The drivers'/driver's passengers was/were hungry again after the meal.})
\ex M\"{u}hendislerin/M\"{u}hendisin kap{\i}c{\i}s{\i} erken \"{o}demeden dolay{\i} sevindiler/sevindi.\\*
({\it The engineers'/engineer's concierge rejoiced\textsubscript{PL/SG} at early payment.})
\ex Pazarc{\i}lar{\i}n/Pazarc{\i}n{\i}n nakliyecisi mesaiden hemen sonra uzand{\i}lar/uzand{\i}.\\*
({\it The marketers'/marketer's carrier lay\textsubscript{PL/SG} down right after work.})
\ex Oyuncular{\i}n/Oyuncunun e\u{g}itimcisi ilk denemede epey zorland{\i}lar/zorland{\i}.\\*
({\it The players'/player's trainer had\textsubscript{PL/SG} a hard time on the first try.})
\ex Zenginlerin/Zenginin animat\"{o}r\"{u} ge\c{c} bir vakitte kalkt{\i}lar/kalkt{\i}.\\*
({\it The rich people's/ric person's animator got\textsubscript{PL/SG} up late.})
\ex Vekillerin/Vekilin polisi m\"{u}thi\c{s} bir a\u{g}r{\i}yla uyand{\i}lar/uyand{\i}.\\*
({\it The deputies'/deputy's police woke\textsubscript{PL/SG} up in terrible pain.})
\ex Par\c{s}\"{o}menlerin/Par\c{s}\"{o}menin h{\i}rs{\i}z{\i} bo\c{s} bir caddede y\"{u}r\"{u}d\"{u}ler/y\"{u}r\"{u}d\"{u}.\\*
({\it The scrolls'/scroll's thief walked\textsubscript{PL/SG} down an empty street.})
\ex Topluluklar{\i}n/Toplulu\u{g}un kral{\i} toplant{\i}dan sonra birden sarard{\i}lar/sarard{\i}.\\*
({\it The communities'/community's king suddenly turned\textsubscript{PL/SG} pale after the meeting.})
\ex \"{U}nl\"{u}lerin/\"{U}nl\"{u}n\"{u}n falc{\i}s{\i} yabanc{\i} bir \"{u}lkede kayboldular/kayboldu.\\*
({\it The famous people's/famous person's fortune teller got\textsubscript{PL/SG} lost in a foreign country.})
\ex \c{C}ift\c{c}ilerin/\c{C}ift\c{c}inin bek\c{c}isi normalden \c{c}ok yava\c{s} gezindiler/gezindi.\\*
({\it The farmers'/farmer's guard roamed\textsubscript{PL/SG} more slowly than usual.})
\z

\section{Filler items}
The filler items used in this experiment were the same as those in Experiment 2A. 

\newpage 
