
\chapter{EXPERIMENT 3 ITEMS} \label{ap:exp3items}

\section{Experimental items}
The experimental items used in this experiment were the same as those in Experiment 2A. 

\section{Filler items}

\subsection{Grammatical fillers}
\ea Profes\"{o}r\"{u}n \"{o}\u{g}rencisi ba\c{s}ar{\i}l{\i} olunca mutlu oldular. \\*
({\it They were happy when the professor's student succeeded.})
\ex \c{C}ocu\u{g}un kedisi d\"{u}n kaybolunca \c{c}{\i}lg{\i}na d\"{o}nd\"{u}ler. \\*
({\it They went crazy when the kid's cat disappeared yesterday.})
\ex Ekibin teknisyeni h{\i}zl{\i} \c{c}al{\i}\c{s}t{\i}\u{g}{\i}ndan tekrar \c{c}a\u{g}{\i}rd{\i}lar. \\*
({\it They called again because the team's technician was working fast.})
\ex Patronun yeme\u{g}i yere d\"{o}k\"{u}l\"{u}nce yenisini yapt{\i}lar. \\*
({\it When the boss' food spilled on the floor, they made a new one.})
\ex Y\"{o}netmenin filmi k\"{o}t\"{u} \c{c}{\i}k{\i}nca dizi izlediler. \\*
({\it When the director's movie turned out to be bad, they watched the series.})
\ex Muhtar{\i}n a\u{g}ac{\i} meyve verince epey \c{s}a\c{s}{\i}rd{\i}lar. \\*
({\it They were quite surprised when the mukhtar's tree bore fruit.})
\ex K\"{o}y\"{u}n \"{o}\u{g}retmeni emekli olunca saatlerce a\u{g}lam{\i}\c{s}lar. \\*
({\it When the village teacher retired, they cried for hours.})
\ex Yaral{\i}n{\i}n ilac{\i} rahats{\i}z edince doktorla konu\c{s}mu\c{s}lar. \\*
({\it When the patient's medicine bothered him, they talked to the doctor.})
\ex M\"{u}\c{s}terinin yeme\u{g}i so\u{g}uk gelince geri g\"{o}nderdiler. \\*
({\it When the customer's food was cold, they sent it back.})
\ex Grubun menajeri onlar{\i} duyunca biraz gerildiler. \\*
({\it They got a little bit nervous when the group's manager heard them.})
\ex Mahallenin \"{u}nl\"{u}s\"{u} kafeye gelince \c{s}ok olmu\c{s}lar. \\*
({\it They were shocked when the celebrity of the neighborhood came to the cafe.})
\ex Tamircinin arabas{\i} sorun \c{c}{\i}kar{\i}nca hemen indiler. \\*
({\it When the mechanic's car got into trouble, they got off right away.})
\ex \c{S}airin \c{s}iiri seyirciler taraf{\i}ndan be\u{g}enilmeyince \"{u}z\"{u}ld\"{u}ler. \\*
({\it They were upset when the poet's poem was not liked by the audience.})
\ex Dizinin ba\c{s}rol\"{u} yanlar{\i}na gelince a\c{s}{\i}r{\i} heyecanland{\i}lar. \\*
({\it They got super excited when the main character of the show came to their side.})
\ex Zanl{\i}n{\i}n avukat{\i} gelince rahat bir nefes ald{\i}lar. \\*
({\it They breathed a sigh of relief when the suspect's lawyer arrived.})
\ex Valinin \c{c}ocu\u{g}u kayboldu\u{g}u i\c{c}in olduk\c{c}a endi\c{s}elenmi\c{s}ler. \\*
({\it They were very worried about the governor's child missing.})
\ex K\"{o}yl\"{u}n\"{u}n elmas{\i} kurtland{\i}\u{g}{\i} i\c{c}in atmak zorunda kald{\i}lar. \\*
({\it The villager's apple was wormed so they had to throw it away.})
\ex Heykeltra\c{s}{\i}n vazosu kolayca yap{\i}\c{s}t{\i}r{\i}l{\i}nca yenisini almad{\i}lar. \\*
({\it The sculptor's vase was easily glued, so they didn't get a new one.})
\ex Tiyatrocunun sa\c{c}{\i} ya\u{g}murda {\i}slan{\i}nca kurulamaya giri\c{s}tiler. \\*
({\it When the actor's hair got wet in the rain, they started to dry it.})
\ex Kom\c{s}unun bebe\u{g}i \"{u}stlerine kusunca banyoya ko\c{s}tular. \\*
({\it They ran to the bathroom when the neighbor's baby vomited on them.})
\ex Sekizinci s{\i}n{\i}flar \"{u}lke ortalamas{\i}n{\i}n \c{c}ok \"{u}st\"{u}ndeler. \\*
({\it Eighth graders are well above the national average.})
\ex \c{C}ocuklar yeni filmi bensiz izlemeye karar verdiler. \\*
({\it The children decided to watch the new movie without me.})
\ex Bizi vazge\c{c}irmek i\c{c}in yola bubi tuza\u{g}{\i} kurabilirler. \\*
({\it They can booby-trap the road to discourage us.})
\ex Yetkililer havaliman{\i}nda daha etkili bir yakla\c{s}{\i}m benimsediler. \\*
({\it Authorities took a more effective approach at the airport.})
\ex Valiler b\"{o}lge meclislerinin se\c{c}imini hep beraber belirlerler. \\*
({\it Governors jointly determined the election of regional councils.})
\ex Asl{\i}nda evde kimse yokken bu kadar \c{s}{\i}marmazlar. \\*
({\it Actually, they don't get so spoiled when nobody's home.})
\ex Insanlar do\u{g}ay{\i} mahvederek kendilerine zarar verdiler. \\*
({\it People have harmed themselves by destroying nature.})
\ex Avrupal{\i}lar Hindistan'{\i} ararken yanl{\i}\c{s}l{\i}kla Amerika'y{\i} ke\c{s}fettiler. \\*
({\it Europeans accidentally discovered Americas while searching for India.})
\ex Bu y\"{o}redeki megalitler tap{\i}nak in\c{s}aas{\i} i\c{c}in ta\c{s}{\i}nm{\i}\c{s} olabilirler. \\*
({\it Megaliths in this area may have been moved for temple construction.})
\ex M\"{u}ttefikler askeri a\c{c}{\i}dan \"{o}nemli avantajlara sahiptirler. \\*
({\it Allies have significant military advantages.})
\z

\subsection{Ungrammatical fillers}
\ea * Bakan{\i}n yard{\i}mc{\i}s{\i} bulununca koltuk geri getirdi. \\*
({\it * When the deputy minister was found, the chair brought back.})
\ex * \"{O}\u{g}rencinin hocas{\i} ayr{\i}l{\i}nca proje birden unuttu. \\*
({\it * The project suddenly forgot when the student's teacher left.})
\ex * Pizzac{\i}n{\i}n kuryesi t\"{o}kezleyince soslar yere sa\c{c}t{\i}. \\*
({\it * When the pizzeria's courier stumbled, sauces spilled onto the floor.})
\ex * Kral{\i}n soytar{\i}s{\i} as{\i}l{\i}nca \c{s}apka yerinde buldu. \\*
({\it * When the king's jester was hanged, the hat found in its place.})
\ex * Dekan{\i}n davetlisi hap\c{s}urunca \c{c}aylar aniden d\"{u}\c{s}\"{u}rd\"{u}. \\*
({\it * The dean's guest sneezed, and the tea suddenly dropped.})
\ex * Dedektifin g\"{o}zl\"{u}k\c{c}\"{u}s\"{u} evlenince hediyeler a\u{g}lanarak verdi. \\*
({\it * When the detective's optician got married, presents gave crying.})
\ex * Politikac{\i}n{\i}n s\"{o}zc\"{u}s\"{u} yakalan{\i}nca a\c{c}{\i}klama haliyle kesti. \\*
({\it * When the politician's spokesperson was caught, the statement cut off.})
\ex * Kad{\i}n{\i}n temizlik\c{c}isi bay{\i}l{\i}nca deterjan tekrar sa\c{c}t{\i}. \\*
({\it * When the woman's housekeeper fainted, the detergent spilled out again.})
\ex * Mankenin ni\c{s}anl{\i}s{\i} vurulunca haber h{\i}zl{\i}ca yayd{\i}. \\*
({\it * News spread quickly when the model's fiancee was shot.})
\ex * \c{C}oban{\i}n s\"{o}zl\"{u}s\"{u} tutuklan{\i}nca kamera sessizce s\"{o}kt\"{u}. \\*
({\it * The camera went silent when the shepherd's spokesman was arrested.})
\ex * Dans\"{o}z\"{u}n kocas{\i} var{\i}nca kap{\i} sakince a\c{c}t{\i}. \\*
({\it * The door opened calmly when the dancer's husband arrived.})
\ex * \c{C}evirmenin kaynanas{\i} aramay{\i}nca metin keyfince bitirdi. \\*
({\it * When the translator's mother-in-law did not call, the text ended arbitrarily.})
\ex * Fabrikat\"{o}r\"{u}n muhasebecisi kovulunca hesap tamamen kar{\i}\c{s}t{\i}rd{\i}. \\*
({\it * The account completely messed up when the fabricator's accountant was fired.})
\ex * \"{U}nl\"{u}n\"{u}n k\"{u}rk\c{c}\"{u}s\"{u} d\"{o}n\"{u}nce kuma\c{s} erkenden dikti. \\*
({\it * When the famous furrier returned, the fabric sewed early.})
\ex * Rekt\"{o}r\"{u}n yard{\i}mc{\i}s{\i} atan{\i}nca k\"{u}t\"{u}phane gece kapatt{\i}. \\*
({\it * The library closed at night when the vice-chancellor was appointed.})
\ex * \c{S}ark{\i}c{\i}n{\i}n taksicisi gecikince trafik aniden kilitledi. \\*
({\it * When the singer's taxi driver was late, traffic suddenly locked up.})
\ex * \c{C}ocu\u{g}un dad{\i}s{\i} aramay{\i}nca bula\c{s}{\i}k saatlerce y{\i}kad{\i}. \\*
({\it * The dished washed for hours when the boy's nanny didn't call.})
\ex * \c{C}ift\c{c}inin tesisat\c{c}{\i}s{\i} gelince borular g\"{u}\c{c}l\"{u}kle s\"{o}kt\"{u}. \\*
({\it * When the farmer's plumber came, the pipes removed with difficulty.})
\ex * \c{C}iftin mobilyac{\i}s{\i} k{\i}z{\i}nca koltuk sinirle par\c{c}alad{\i}. \\*
({\it * When the couple's furniture maker got angry, the sofa smashed in anger.})
\ex * Adam{\i}n falc{\i}s{\i} yan{\i}l{\i}nca fincan \"{o}fkeyle k{\i}rd{\i}. \\*
({\it * When the man's fortune teller was wrong, the cup broke in anger.})
\ex * Polisin eve gelmesi \c{c}ocu\u{g}a ans{\i}z{\i}n s\"{o}yledi. \\*
({\it * The fact that the police will come to house said to the child abruptly.})
\ex * Kahvenin para \"{o}demeyince barista arkas{\i}ndan ko\c{s}tu. \\*
({\it * When he did not paid money coffee, the barista ran after him.})
\ex * Amerika vergiyi reddedince ilk \c{c}{\i}kan sava\c{s} kazand{\i}. \\*
({\it * When America refused the pay taxes, the war won.})
\ex * Film sanat\c{c}{\i}n{\i}n aile nefret etmesini anlat{\i}yor. \\*
({\it * The movie explains the hatred the artist had family.})
\ex * Restoran sakinleri bina giri\c{s}i y\"{o}nlendirmeyi unutmu\c{s}. \\*
({\it * Restaurant residents forgot to redirect entering into the building.})
\ex * Usta bir gitarist oldu\u{g}u hayran olundu. \\*
({\it * He was loved by many he is a great guitarist.})
\ex * Trafik {\i}\c{s}{\i}klar{\i} s\"{u}r\"{u}c\"{u}n\"{u}n kafas{\i}n{\i} kar{\i}\c{s}t{\i}rmaya denedi. \\*
({\it * Traffic lights tried by confuse the driver.})
\ex * Ev arkada\c{s}{\i} belki de birini kesin g\"{o}rm\"{u}\c{s}. \\*
({\it * The roommate may have absolutely seen someone for sure.})
\ex * Hastaneye var{\i}nca doktor hastadan aray{\i}p durmu\c{s}. \\*
({\it * Arriving at the hospital, the doctor kept calling from the patient.})
\ex * Yenisine erik re\c{c}eli bitti\u{g}inden dolay{\i} aranm{\i}\c{s}. \\*
({\it *  A new one had was searched for due to running out of plum jam.})
\z

\newpage 