
\chapter{EXPERIMENT 1 ITEMS} \label{ap:exp1items}

\section{Experimental items}

\ea Y\"{o}neticilerin/Y\"{o}neticinin a\c{s}\c{c}{\i}s{\i} mutfakta s\"{u}rekli z{\i}plad{\i}lar/z{\i}plad{\i}. \\* 
({\it The managers'/manager's cook jumped\textsubscript{PL/SG} constantly in the kitchen.})
\ex \"{O}\u{g}rencilerin/\"{O}\u{g}rencinin ablas{\i} s{\i}n{\i}fta birden bay{\i}ld{\i}lar/bay{\i}ld{\i}. \\* 
({\it The students'/student's older sister suddenly fainted\textsubscript{PL/SG} in class.})
\ex Marangozlar{\i}n/Marangozun abisi at\"{o}lyeden h{\i}zla uzakla\c{s}t{\i}lar/uzakla\c{s}t{\i}.\\* 
({\it The carpenters'/carpenter's older brother rushed\textsubscript{PL/SG} away from the workshop.})
\ex Mahallelilerin/Mahallelinin emlak\c{c}{\i}s{\i} aniden k\"{u}stah\c{c}a g\"{u}ld\"{u}ler/g\"{u}ld\"{u}.\\* 
({\it The residents'/resident's real estate agent suddenly laughed\textsubscript{PL/SG} arrogantly.})
\ex K{\i}zlar{\i}n/K{\i}z{\i}n halas{\i} sabaha kadar a\u{g}lad{\i}lar/a\u{g}lad{\i}.\\* 
({\it The girls'/girl's aunt cried\textsubscript{PL/SG} until morning.})
\ex Damatlar{\i}n/Damat{\i}n day{\i}s{\i} arada s{\i}rada s{\i}k{\i}ld{\i}lar/s{\i}k{\i}ld{\i}.\\* 
({\it Grooms'/Groom's uncle are bored\textsubscript{PL/SG} from time to time.})
\ex Doktorlar{\i}n/Doktorun \c{c}i\c{c}ek\c{c}isi g\"{u}nden g\"{u}ne zay{\i}flad{\i}lar/zay{\i}flad{\i}.\\* 
({\it The doctors'/doctor's florist got\textsubscript{PL/SG} thinner day by day.})
\ex Aristokratlar{\i}n/Aristokrat{\i}n hizmet\c{c}isi yorgun arg{\i}n yatt{\i}lar/yatt{\i}.\\*
({\it Aristocrats'/Aristocrat's maid slept\textsubscript{PL/SG} tiredly.})
\ex Konu\c{s}mac{\i}lar{\i}n/Konu\c{s}mac{\i}n{\i}n sunucusu ola\u{g}an\"{u}st\"{u} h{\i}zl{\i} ko\c{s}tular/ko\c{s}tu.\\*
({\it Speakers'/Speaker's presenter ran\textsubscript{PL/SG} extraordinarily fast.})
\ex Psikiyatristlerin/Psikiyatristin eczac{\i}s{\i} a\c{c} susuz dola\c{s}t{\i}lar/dola\c{s}t{\i}.\\*
({\it Psychiatrists'/Psychiatrist's pharmacist wandered\textsubscript{PL/SG} hungry and thirsty.})
\ex Politikac{\i}lar{\i}n/Politikac{\i}n{\i}n hocas{\i} adliyeden \c{c}abucak \c{c}{\i}kt{\i}lar/\c{c}{\i}kt{\i}.\\*
({\it The politicians'/politician's teacher quickly left\textsubscript{PL/SG} the courthouse.})
\ex Hakimlerin/Hakimin \c{c}ayc{\i}s{\i} nedensiz yere k{\i}zd{\i}lar/k{\i}zd{\i}.\\*
({\it Judges'/Judge's tea-maker got\textsubscript{PL/SG} angry for no reason.})
\ex Oyuncular{\i}n/Oyuncunun hem\c{s}iresi etrafta ama\c{c}s{\i}zca gezdiler/gezdi.\\*
({\it Actors'/Actor's nurse wandered\textsubscript{PL/SG} around aimlessly.})
\ex \"{O}\u{g}retmenlerin/\"{O}\u{g}retmenin m\"{u}diresi biraz \"{o}nce arad{\i}lar/arad{\i}.\\*
({\it Teachers'/Teacher's headmistress just called\textsubscript{PL/SG}.})
\ex Milyonerlerin/Milyonerin terzisi tamamen gereksizce ba\u{g}{\i}rd{\i}lar/ba\u{g}{\i}rd{\i}.\\*
({\it The millionaires'/millionaire's tailor yelled\textsubscript{PL/SG} completely unnecessarily.})
\ex Bebeklerin/Bebek\u{g}in bak{\i}c{\i}s{\i} \c{c}ok kibar davrand{\i}lar/davrand{\i}.\\*
({\it Babies'/The baby caretaker was/were very kind.})
\ex \c{C}ocuklar{\i}n/\c{C}ocu\u{g}un dad{\i}s{\i} y\"{u}ksek sesle konu\c{s}tular/konu\c{s}tu.\\*
({\it The children's/child's nanny spoke\textsubscript{PL/SG} loudly.})
\ex Futbolcular{\i}n/Futbolcunun s\"{u}r\"{u}c\"{u}s\"{u} \c{c}ok yava\c{s} \c{c}al{\i}\c{s}t{\i}lar/\c{c}al{\i}\c{s}t{\i}.\\*
({\it The players'/player's driver ran\textsubscript{PL/SG} very slowly.})
\ex Modac{\i}lar{\i}n/Modac{\i}n{\i}n taksicisi saatlerce durmaks{\i}z{\i}n i\c{c}tiler/i\c{c}ti.\\*
({\it The fashionistas'/fashionista's taxi driver drank\textsubscript{PL/SG} non-stop for hours.})
\ex Sanat\c{c}{\i}lar{\i}n/Sanat\c{c}{\i}n{\i}n \c{c}alg{\i}c{\i}s{\i} feci bir \c{s}ekilde \"{o}ld\"{u}ler/\"{o}ld\"{u}.\\*
({\it The artists'/artist's instrumentalist died\textsubscript{PL/SG} tragically.})
\ex Dedektiflerin/Dedektifin di\c{s}\c{c}isi ilk kez \c{c}{\i}lg{\i}nca e\u{g}lendiler/e\u{g}lendi.\\*
({\it The detectives'/detective's dentist had\textsubscript{PL/SG} a blast for the first time.})
\ex Esnaflar{\i}n/Esnaf{\i}n m\"{u}\c{s}terisi \c{s}ikayettten hemen sonra sustular/sustu.\\*
({\it The shopkeepers'/shopkeeper's customer fell\textsubscript{PL/SG} silent immediately after complaining.})
\ex \c{S}ark{\i}c{\i}lar{\i}n/\c{S}ark{\i}c{\i}n{\i}n korumas{\i} her zamanki gibi geciktiler/gecikti.\\*
({\it The singers'/singer's bodyguard was/were delayed as usual.})
\ex G\"{o}stericilerin/G\"{o}stericinin izleyicisi ak\c{s}ama kadar sessizce oturdular/oturdu.\\*
({\it The demonstrators'/demonstrator's audience sat\textsubscript{PL/SG} in silence until the evening.})
\ex Cerrahlar{\i}n/Cerrah{\i}n hastas{\i} ak\c{s}amki g\"{o}steriden \"{o}nce ka\c{c}t{\i}lar/ka\c{c}t{\i}.\\*
({\it The surgeons'/surgeon's patients fled\textsubscript{PL/SG} before the evening's performance.})
\ex Dalg{\i}\c{c}lar{\i}n/Dalg{\i}\c{c}{\i}n annesi bile bile ge\c{c} kald{\i}lar/kald{\i}.\\*
({\it The divers'/diver's mother was/were late on purpose.})
\ex Fabrikat\"{o}rlerin/Fabrikat\"{o}r\"{u}n i\c{s}\c{c}isi beklenmedik bir anda hastaland{\i}lar/hastaland{\i}.\\*
({\it Fabricators'/Fabricator's worker got\textsubscript{PL/SG} sick unexpectedly.})
\ex Komedyenlerin/Komedyenin yard{\i}mc{\i}s{\i} poyrazdan dolay{\i} \"{u}\c{s}\"{u}d\"{u}ler/\"{u}\c{s}\"{u}d\"{u}.\\*
({\it The comedians'/comedian's assistant was/were cold because of the bad wind.})
\ex \c{S}of\"{o}rlerin/\c{S}of\"{o}r\"{u}n yolcusu yemekten sonra yine ac{\i}kt{\i}lar/ac{\i}kt{\i}.\\*
({\it Drivers'/Driver's passenger got\textsubscript{PL/SG} hungry again after dinner.})
\ex M\"{u}hendislerin/M\"{u}hendisin kap{\i}c{\i}s{\i} erken \"{o}demeden dolay{\i} sevindiler/sevindi.\\*
({\it Engineers'/Engineer's doorman delighted\textsubscript{PL/SG} with early payment.})
\ex Pazarc{\i}lar{\i}n/Pazarc{\i}n{\i}n nakliyecisi mesaiden hemen sonra uzand{\i}lar/uzand{\i}.\\*
({\it The marketers'/marketer's shipper lay\textsubscript{PL/SG} down right after work.})
\ex Oyuncular{\i}n/Oyuncunun e\u{g}itimcisi ilk denemede epey zorland{\i}lar/zorland{\i}.\\*
({\it Players'/Player's trainer struggled\textsubscript{PL/SG} on the first try.})
\ex Mankenlerin/Mankenin modac{\i}s{\i} ge\c{c} bir vakitte kalkt{\i}lar/kalkt{\i}.\\*
({\it The mannequins'/The mannequin's fashionista got\textsubscript{PL/SG} up late.})
\ex Konuklar{\i}n/Konu\u{g}un teyzesi m\"{u}thi\c{s} bir a\u{g}r{\i}yla uyand{\i}lar/uyand{\i}.\\*
({\it Guests'/Guest's aunt woke\textsubscript{PL/SG} up with excruciating pain.})
\ex O\u{g}lanlar{\i}n/O\u{g}lan{\i}n amcas{\i} bo\c{s} bir caddede y\"{u}r\"{u}d\"{u}ler/y\"{u}r\"{u}d\"{u}.\\*
({\it The boys'/boy's uncle walked\textsubscript{PL/SG} down an empty street.})
\ex Avukatlar{\i}n/Avukat{\i}n kom\c{s}usu toplant{\i}dan sonra birden sarard{\i}lar/sarard{\i}.\\*
({\it The lawyers'/lawyer's neighbor suddenly turned\textsubscript{PL/SG} yellow after the meeting.})
\ex \"{U}nl\"{u}lerin/\"{U}nl\"{u}n\"{u}n falc{\i}s{\i} yabanc{\i} bir \"{u}lkede kayboldular/kayboldu.\\*
({\it Celebrities'/Celebrity's fortune teller got\textsubscript{PL/SG} lost in a foreign country.})
\ex \c{C}ift\c{c}ilerin/\c{C}ift\c{c}inin bek\c{c}isi normalden \c{c}ok yava\c{s} gezindiler/gezindi.\\*
({\it The Farmers'/Farmer's guard roamed\textsubscript{PL/SG} a lot slower than usual.})
\ex Kad{\i}nlar{\i}n/Kad{\i}n{\i}n ninesi ge\c{c}en seneye g\"{o}re din\c{c}le\c{s}tiler/din\c{c}le\c{s}ti.\\*
({\it Women's/Woman's grandmother became\textsubscript{PL/SG} more vigorous compared to last year.})
\z

\section{Filler items}
\ea {Grammatical Fillers}
\ea Adam{\i}n annesi fenala\c{s}{\i}nca inek kurban ettiler.\\*
({\it When the man's mother fell ill, they sacrificed a cow.})
\ex Sosyologun \"{o}\u{g}rencisi konu\c{s}unca tutars{\i}zl{\i}k a\c{c}{\i}\u{g}a \c{c}{\i}kard{\i}lar.\\*
({\it When the sociologist's student spoke, they revealed inconsistency.})
\ex Doktorun hem\c{s}iresi gelene kadar hasta taburcu ettiler.\\*
({\it The patient was discharged until the doctor's nurse arrived.})
\ex Kemanc{\i}n{\i}n sevgilisi \"{o}l\"{u}nce mezar ziyaret ettiler.\\*
({\it When the violinist's lover died, they visited a grave.})
\ex Hocan{\i}n kap{\i}c{\i}s{\i} bay{\i}l{\i}nca doktor rahats{\i}z ettiler.\\*
({\it When the teacher's doorman fainted, they bothered a doctor.})
\ex Medyumun kocas{\i} sa\c{c}malay{\i}nca falc{\i} zengin ettiler.\\*
({\it When the psychic's husband made nonsense, they made the fortune-teller rich.})
\ex Ba\c{s}kan{\i}n di\c{s}\c{c}isi t{\i}rs{\i}nca stajyer kabul ettiler.\\*
({\it When the president's dentist got scared, they accepted a trainee.})
\ex Ele\c{s}tirmenin kar{\i}s{\i} k{\i}v{\i}rt{\i}nca sap{\i}k tahrik ettiler.\\*
({\it Perverted aroused when the critic's wife squirmed.})
\ex Patronun kahyas{\i} d\"{u}\c{s}\"{u}nce d\"{u}\c{s}man mutlu ettiler.\\*
({\it When the boss's butler fell, they made an enemy happy.})
\ex M\"{u}d\"{u}r\"{u}n a\c{s}\c{c}{\i}s{\i} haz{\i}rlan{\i}nca yemek haz{\i}r ettiler.\\*
({\it When the manager's cook was ready, they prepared a meal.})
\ex \c{C}ocu\u{g}un abisi \"{u}z\"{u}l\"{u}nce oyuncak icat ettiler.\\*
({\it When the boy's brother was upset, they invented a toy.})
\ex Psikologun hastas{\i} gecikince vakit hi\c{c} ettiler.\\*
({\it When the psychologist's patient was late, they killed some time.})
\ex Ressam{\i}n tedarik\c{c}isi kaybolunca tuval ithal ettiler.\\*
({\it When the painter's supplier disappeared, they imported canvas.})
\ex Di\c{s}\c{c}inin temizlik\c{c}isi yorulunca hademe ikna ettiler.\\*
({\it When the dentist's housekeeper got tired, they convinced the janitor.})
\ex Kimyagerin kuryesi hap\c{s}urunca deney ak{\i}l ettiler.\\*
({\it When the chemist's courier sneezed, they came up with the experiment.})
\ex Mankenin motorcusu s{\i}z{\i}nca \c{c}{\i}rak me\c{s}gul ettiler.\\*
({\it They kept the apprentice busy when the model's biker fell asleep.})
\ex Dekan{\i}n davetlisi ge\c{c}ince seyirci aya\u{g}a kald{\i}rd{\i}lar.\\*
({\it When the dean's guest passed, they made the audience stood up.})
\ex Mafyan{\i}n yat{\i}r{\i}mc{\i}s{\i} bat{\i}nca kuyumcu tehdit ettiler.\\*
({\it When the investor of the mafia went bankrupt, they threatened the jeweler.})
\ex A\c{s}\c{c}{\i}n{\i}n manav{\i} kapan{\i}nca et tedarik ettiler.\\*
({\it When the cook's grocer closed, they supplied meat.})
\ex \"{O}\u{g}rencinin hocas{\i} anlat{\i}nca makine icat ettiler.\\*
({\it When the student's teacher told them, they invented a machine.})
\z
\ex {Ungrammatical Fillers}
\ea * Bakan{\i}n yard{\i}mc{\i}s{\i} bulununca koltuk geri getirdi.\\*
({\it * When the deputy minister was found, the chair brought back.})
\ex * \"{O}\u{g}rencinin hocas{\i} ayr{\i}l{\i}nca proje birden unuttu.\\*
({\it * The project was suddenly forgot when the student's teacher left.})
\ex * Pizzac{\i}n{\i}n kuryesi t\"{o}kezleyince soslar yere sa\c{c}t{\i}.\\*
({\it *  When the pizzeria's courier stumbled, sauces spilled onto the floor.})
\ex * Kral{\i}n soytar{\i}s{\i} as{\i}l{\i}nca \c{s}apka yerinde buldu.\\*
({\it * When the king's jester was hanged, the hat found in its place.})
\ex * Dekan{\i}n davetlisi hap\c{s}urunca \c{c}aylar aniden d\"{u}\c{s}\"{u}rd\"{u}.\\*
({\it * The dean's guest sneezed, and the tea suddenly dropped.})
\ex * Dedektifin g\"{o}zl\"{u}k\c{c}\"{u}s\"{u} evlenince hediyeler a\u{g}lanarak verdi.\\*
({\it * When the detective's optician got married, he gave presents crying.})
\ex * Politikac{\i}n{\i}n s\"{o}zc\"{u}s\"{u} yakalan{\i}nca a\c{c}{\i}klama haliyle kesti.\\*
({\it * When the politician's spokesperson was caught, the statement naturally cut off.})
\ex * Kad{\i}n{\i}n temizlik\c{c}isi bay{\i}l{\i}nca deterjan tekrar sa\c{c}t{\i}.\\*
({\it * When the woman's housekeeper fainted, the detergent spilled out again.})
\ex * Mankenin ni\c{s}anl{\i}s{\i} vurulunca haber h{\i}zl{\i}ca yayd{\i}.\\*
({\it * News spread quickly when the model's fiancee was shot.})
\ex * \c{C}oban{\i}n s\"{o}zl\"{u}s\"{u} tutuklan{\i}nca kamera sessizce s\"{o}kt\"{u}.\\*
({\it * The camera discharge when the shepherd's spokesman was arrested.})
\ex * Dans\"{o}z\"{u}n kocas{\i} var{\i}nca kap{\i} sakince a\c{c}t{\i}.\\*
({\it * The door opened calmly when the dancer's husband arrived.})
\ex * \c{C}evirmenin kaynanas{\i} aramay{\i}nca metin keyfince bitirdi.\\*
({\it * When the translator's mother-in-law did not call, the text finished arbitrarily.})
\ex * Fabrikat\"{o}r\"{u}n muhasebecisi kovulunca hesap tamamen kar{\i}\c{s}t{\i}rd{\i}.\\*
({\it * The account completely messed up when the fabricator's accountant was fired.})
\ex * \"{U}nl\"{u}n\"{u}n k\"{u}rk\c{c}\"{u}s\"{u} d\"{o}n\"{u}nce kuma\c{s} erkenden dikti.\\*
({\it * When the famous furrier returned, the fabric sewed early.})
\ex * Rekt\"{o}r\"{u}n yard{\i}mc{\i}s{\i} atan{\i}nca k\"{u}t\"{u}phane gece kapatt{\i}.\\*
({\it * The library closed at night when the vice-chancellor was appointed.})
\ex * \c{S}ark{\i}c{\i}n{\i}n taksicisi gecikince trafik aniden kilitledi.\\*
({\it * When the singer's taxi driver was late, traffic suddenly locked.})
\ex * \c{C}ocu\u{g}un dad{\i}s{\i} aramay{\i}nca bula\c{s}{\i}k saatlerce y{\i}kad{\i}.\\*
({\it * The dishes cleanerd for hours when the boy's nanny didn't call.})
\ex * \c{C}ift\c{c}inin tesisat\c{c}{\i}s{\i} gelince borular g\"{u}\c{c}l\"{u}kle s\"{o}kt\"{u}.\\*
({\it * When the farmer's plumber came, the pipes had to remove with difficulty.})
\ex * \c{C}iftin mobilyac{\i}s{\i} k{\i}z{\i}nca koltuk sinirle par\c{c}alad{\i}.\\*
({\it * When the couple's furniture maker got angry, the sofa smashed in anger.})
\ex * Adam{\i}n falc{\i}s{\i} yan{\i}l{\i}nca fincan \"{o}fkeyle k{\i}rd{\i}.\\*
({\it * When the man's fortune teller was wrong, the cup broke in anger.})
\z
\z


\newpage
