
\chapter{EXPERIMENT 2A ITEMS} \label{ap:exp2aitems}
\section{Experimental items}

\ea D\"{o}vd\"{u}kleri/D\"{o}vd\"{u}\u{g}\"{u} \c{c}ocuk okula yorgun arg{\i}n geldiler/geldi.\\*
({\it The boy they/he beaten came\textsubscript{PL/SG} to school exhausted.})
\ex Tuttuklar{\i}/Tuttu\u{g}u a\c{s}\c{c}{\i} mutfakta s\"{u}rekli z{\i}plad{\i}lar/z{\i}plad{\i}.\\*
({\it The cook they/he hired bounced\textsubscript{PL/SG} around in the kitchen.})
\ex Tan{\i}d{\i}klar{\i}/Tan{\i}d{\i}\u{g}{\i} m\"{u}d\"{u}r s{\i}n{\i}fta birden bay{\i}ld{\i}lar/bay{\i}ld{\i}.\\*
({\it The principal they/he knew fainted\textsubscript{PL/SG} in class.})
\ex G\"{o}rd\"{u}kleri/G\"{o}rd\"{u}\u{g}\"{u} marangoz at\"{o}lyeden h{\i}zla uzakla\c{s}t{\i}lar/uzakla\c{s}t{\i}.\\*
({\it The carpenter they/he saw hurried\textsubscript{PL/SG} away from the workshop.})
\ex Azarlad{\i}klar{\i}/Azarlad{\i}\u{g}{\i} emlak\c{c}{\i} aniden k\"{u}stah\c{c}a g\"{u}ld\"{u}ler/g\"{u}ld\"{u}.\\*
({\it The realtor they/he scolded suddenly laughed\textsubscript{PL/SG} arrogantly.})
\ex Reddetikleri/Reddeti\u{g}i akademisyen sabaha kadar a\u{g}lad{\i}lar/a\u{g}lad{\i}.\\*
({\it The academic, whom they/he rejected, wept\textsubscript{PL/SG} until the morning.})
\ex Beklettikleri/Beklettik\u{g}i ara\c{s}t{\i}rmac{\i} g\"{u}n boyunca s{\i}k{\i}ld{\i}lar/s{\i}k{\i}ld{\i}.\\*
({\it The investigator they/he were holding was/were bored during the day.})
\ex Bakt{\i}klar{\i}/Bakt{\i}\u{g}{\i} hasta g\"{u}nden g\"{u}ne zay{\i}flad{\i}lar/zay{\i}flad{\i}.\\*
({\it The patient they/he cared for got\textsubscript{PL/SG} weaker day by day.})
\ex Yorduklar{\i}/Yordu\u{g}u oyuncu onikiden \"{o}nce uyudular/uyudu.\\*
({\it The player they/he wore down fell\textsubscript{PL/SG} asleep before twelve.})
\ex \c{C}al{\i}\c{s}t{\i}rd{\i}klar{\i}/\c{C}al{\i}\c{s}t{\i}rd{\i}\u{g}{\i} hizmet\c{c}i yorgun arg{\i}n yatt{\i}lar/yatt{\i}.\\*
({\it The maid they/he hired went\textsubscript{PL/SG} to bed tiredly.})
\ex Kovduklar{\i}/Kovdu\u{g}u sunucu ola\u{g}an\"{u}st\"{u} bir h{\i}zla konu\c{s}tular/konu\c{s}tu.\\*
({\it The server they/he fired spoke\textsubscript{PL/SG} with extraordinary speed.})
\ex Kaybettikleri/Kaybetti\u{g}i turist a\c{c} susuz dola\c{s}t{\i}lar/dola\c{s}t{\i}.\\*
({\it The tourist they/he lost wandered\textsubscript{PL/SG} around hungry and thirsty.})
\ex Cezaland{\i}rd{\i}klar{\i}/Cezaland{\i}rd{\i}\u{g}{\i} hoca hapisten \c{c}abucak \c{c}{\i}kt{\i}lar/\c{c}{\i}kt{\i}.\\*
({\it The teacher they/he punished\textsubscript{PL/SG} quickly got out of jail.})
\ex Uyand{\i}rd{\i}klar{\i}/Uyand{\i}rd{\i}\u{g}{\i} \c{c}ayc{\i} nedensiz yere k{\i}zd{\i}lar/k{\i}zd{\i}.\\*
({\it The tea shop they/he woke\textsubscript{PL/SG} up got angry for no reason.})
\ex Susturduklar{\i}/Susturdu\u{g}u hem\c{s}ire etrafta ama\c{c}s{\i}zca gezdiler/gezdi.\\*
({\it The nurse they/he had silenced wandered\textsubscript{PL/SG} around aimlessly.})
\ex Sorduklar{\i}/Sordu\u{g}u m\"{u}dire biraz \"{o}nce arad{\i}lar/arad{\i}.\\*
({\it The manager they/he asked just called\textsubscript{PL/SG}.})
\ex G\"{o}nderdikleri/G\"{o}nderdi\u{g}i terzi tamamen gereksizce ba\u{g}{\i}rd{\i}lar/ba\u{g}{\i}rd{\i}.\\*
({\it The tailor they/he sent yelled\textsubscript{PL/SG} completely unnecessarily.})
\ex Bulduklar{\i}/Bulduk\u{g}u bak{\i}c{\i} \c{c}ok kibar davrand{\i}lar/davrand{\i}.\\*
({\it The caretaker they/he found was/were very kind.})
\ex Be\u{g}endikleri/Be\u{g}endi\u{g}i dad{\i} sahil boyunca s\"{u}z\"{u}ld\"{u}ler/s\"{u}z\"{u}ld\"{u}.\\*
({\it The nanny they/he liked glided\textsubscript{PL/SG} along the beach.})
\ex Ara\c{s}t{\i}rd{\i}klar{\i}/Ara\c{s}t{\i}rd{\i}\u{g}{\i} tamirci \c{c}ok yava\c{s} \c{c}al{\i}\c{s}t{\i}lar/\c{c}al{\i}\c{s}t{\i}.\\*
({\it The mechanic they/he searched worked\textsubscript{PL/SG} very slowly.})
\ex Efkarland{\i}rd{\i}klar{\i}/Efkarland{\i}rd{\i}\u{g}{\i} taksici saatlerce durmaks{\i}z{\i}n i\c{c}tiler/i\c{c}ti.\\*
({\it The taxi driver they/he made nostalgic drank\textsubscript{PL/SG} nonstop for hours.})
\ex Kovalad{\i}klar{\i}/Kovalad{\i}\u{g}{\i} \c{c}alg{\i}c{\i} feci bir \c{s}ekilde \"{o}ld\"{u}ler/\"{o}ld\"{u}.\\*
({\it The instrumentalist they/he chased died\textsubscript{PL/SG} horribly.})
\ex Gittikleri/Gitti\u{g}i di\c{s}\c{c}i ilk kez \c{c}{\i}lg{\i}nca e\u{g}lendiler/e\u{g}lendi.\\*
({\it The dentist they/he went to had\textsubscript{PL/SG} fun for the first time.})
\ex A\u{g}latt{\i}klar{\i}/A\u{g}latt{\i}\u{g}{\i} m\"{u}\c{s}teri \c{s}ikayetinden hemen sonra sustular/sustu.\\*
({\it The customer they/he made cry shut\textsubscript{PL/SG} up immediately after his complaint.})
\ex \c{C}{\i}ld{\i}rtt{\i}klar{\i}/\c{C}{\i}ld{\i}rtt{\i}\u{g}{\i} koruma her zamanki gibi geciktiler/gecikti.\\*
({\it The bodyguard they/he drove crazy delayed\textsubscript{PL/SG} as usual.})
\ex Getirdikleri/Getirdi\u{g}i izleyici ak\c{s}ama kadar sessizce oturdular/oturdu.\\*
({\it The audience they/he brought sat\textsubscript{PL/SG} in silence until the evening.})
\ex Delirttikleri/Delirtti\u{g}i hasta ak\c{s}amki muayeneden \"{o}nce ka\c{c}t{\i}lar/ka\c{c}t{\i}.\\*
({\it The patient they/he had driven mad fled\textsubscript{PL/SG} before the evening's examination.})
\ex Anla\c{s}t{\i}klar{\i}/Anla\c{s}t{\i}\u{g}{\i} terapist bile bile ge\c{c} kald{\i}lar/kald{\i}.\\*
({\it The therapist they/he hired was/were late on purpose.})
\ex G\"{u}vendikleri/G\"{u}vendi\u{f}i i\c{s}\c{c}i beklenmedik bir anda hastaland{\i}lar/hastaland{\i}.\\*
({\it The employee they/he trusted fell\textsubscript{PL/SG} ill unexpectedly.})
\ex E\u{g}ittikleri/E\u{g}itti\u{g}i hostes sert r\"{u}zgarlardan dolay{\i} \"{u}\c{s}\"{u}d\"{u}ler/\"{u}\c{s}\"{u}d\"{u}.\\*
({\it The stewardess they/he trained got\textsubscript{PL/SG} cold from the strong winds.})
\ex Doyurduklar{\i}/Doyurdu\u{g}u yolcu yemekten sonra yine ac{\i}kt{\i}lar/ac{\i}kt{\i}.\\*
({\it The passengers they/he fed got\textsubscript{PL/SG} hungry again after the meal.})
\ex \c{C}a\u{g}{\i}rd{\i}klar{\i}/\c{C}a\u{g}{\i}rd{\i}\u{g}{\i} kap{\i}c{\i} erken \"{o}demeden dolay{\i} sevindiler/sevindi.\\*
({\it The concierge they/he called was/were happy with the early payment.})
\ex Yorduklar{\i}/Yordu\u{g}u nakliyeci mesaiden hemen sonra uzand{\i}lar/uzand{\i}.\\*
({\it The shipper they/he had tired of lay\textsubscript{PL/SG} down right after work.})
\ex Yeti\c{s}tirdikleri/Yeti\c{s}tirdi\u{g}i e\u{g}itimci ilk denemede epey zorland{\i}lar/zorland{\i}.\\*
({\it The trainer they/he trained had\textsubscript{PL/SG} a hard time on the first try.})
\ex Kiralad{\i}klar{\i}/Kiralad{\i}\u{g}{\i} animat\"{o}r ge\c{c} bir vakitte kalkt{\i}lar/kalkt{\i}.\\*
({\it The animator they/he hired got\textsubscript{PL/SG} up late.})
\ex Yaralad{\i}klar{\i}/Yaralad{\i}\u{g}{\i} polis m\"{u}thi\c{s} bir a\u{g}r{\i}yla uyand{\i}lar/uyand{\i}.\\*
({\it The cop they/he had injured woke\textsubscript{PL/SG} up in terrible pain.})
\ex Ka\c{c}{\i}rd{\i}klar{\i}/Ka\c{c}{\i}rd{\i}\u{g}{\i} h{\i}rs{\i}z bo\c{s} bir caddede y\"{u}r\"{u}d\"{u}ler/y\"{u}r\"{u}d\"{u}.\\*
({\it The thief they/he kidnapped walked\textsubscript{PL/SG} down an empty street.})
\ex Zehirledikleri/Zehirledi\u{g}i kral toplant{\i}dan sonra birden sarard{\i}lar/sarard{\i}.\\*
({\it The king they/he poisoned turned\textsubscript{PL/SG} pale after the meeting.})
\ex Gezdirdikleri/Gezdirdi\u{g}i falc{\i} yabanc{\i} bir \"{u}lkede kayboldular/kayboldu.\\*
({\it The fortune teller they/he showed around got\textsubscript{PL/SG} lost in a foreign country.})
\ex \c{S}\"{u}phelendirdikleri/\c{S}\"{u}phelendirdi\u{g}i bek\c{c}i normalden \c{c}ok yava\c{s} gezindiler/gezindi.\\*
({\it The guard they/he suspected roamed\textsubscript{PL/SG} much slower than normal.})
\z

\section{Filler items}
\ea {Grammatical Fillers}
\ea Okuttuklar{\i} \"{o}\u{g}renci ba\c{s}ar{\i}l{\i} olunca mutlu oldular. \\*
({\it They were happy when the student they sponsored was successful.})
\ex Biriktirdikleri para d\"{u}n kaybolunca \c{c}{\i}lg{\i}na d\"{o}nd\"{u}ler. \\*
({\it They went crazy when their savings disappeared yesterday.})
\ex D\"{u}\c{s}\"{u}nd\"{u}kleri teknisyen h{\i}zl{\i} \c{c}al{\i}\c{s}t{\i}\u{g}{\i}ndan tekrar \c{c}a\u{g}{\i}rd{\i}lar. \\*
({\it They called again because the technician they thought about was working fast.})
\ex Haz{\i}rlad{\i}klar{\i} yemek yere d\"{o}k\"{u}l\"{u}nce yenisini yapt{\i}lar. \\*
({\it When the food they prepared spilled on the floor, they made a new one.})
\ex Ba\c{s}lad{\i}klar{\i} film k\"{o}t\"{u} \c{c}{\i}k{\i}nca dizi izlediler. \\*
({\it When the movie they started watching was bad, they watched a TV series.})
\ex Diktikleri a\u{g}a\c{c} meyve verince epey \c{s}a\c{s}{\i}rd{\i}lar. \\*
({\it They were quite surprised when the tree they planted bore fruit.})
\ex Sevdikleri \"{o}\u{g}retmen emekli olunca saatlerce a\u{g}lam{\i}\c{s}lar. \\*
({\it They cried for hours when their favorite teacher retired.})
\ex Kulland{\i}klar{\i} ila\c{c} rahats{\i}z edince doktorla konu\c{s}mu\c{s}lar. \\*
({\it They talked to the doctor when the medication they used bothered them.})
\ex S\"{o}yledikleri yemek so\u{g}uk gelince geri g\"{o}nderdiler. \\*
({\it When the food they ordered was cold, they sent it back.})
\ex Ay{\i}plad{\i}klar{\i} kad{\i}n onlar{\i} duyunca biraz gerildiler. \\*
({\it When the woman they blamed heard them, they got a little nervous.})
\ex Bahsettikleri \"{u}nl\"{u} kafeye gelince \c{s}ok olmu\c{s}lar. \\*
({\it They were shocked when the famous person they were talking about came to the cafe.})
\ex Bindikleri araba sorun \c{c}{\i}kar{\i}nca hemen indiler. \\*
({\it When the car they were in caused trouble, they got off immediately.})
\ex Okuduklar{\i} \c{s}iir seyirciler taraf{\i}ndan be\u{g}enilmeyince \"{u}z\"{u}ld\"{u}ler. \\*
({\it They were upset when the poem they read was not liked by the audience.})
\ex Dinledikleri \c{s}ark{\i}c{\i} yanlar{\i}na gelince a\c{s}{\i}r{\i} heyecanland{\i}lar. \\*
({\it They got very excited when the singer they were listening to came to them.})
\ex Ka\c{c}t{\i}klar{\i} katil durunca rahat bir nefes ald{\i}lar. \\*
({\it They breathed a sigh of relief when the killer they were running from stopped.})
\ex Kar\c{s}{\i}la\c{s}t{\i}klar{\i} \c{c}ocuk kayboldu\u{g}u i\c{c}in olduk\c{c}a endi\c{s}elenmi\c{s}ler. \\*
({\it They were very worried that the boy they met had disappeared.})
\ex Ald{\i}klar{\i} elma kurtland{\i}\u{g}{\i} i\c{c}in atmak zorunda kald{\i}lar. \\*
({\it The apple they bought got wormy so they had to throw it away.})
\ex K{\i}rd{\i}klar{\i} tabak kolayca yap{\i}\c{s}t{\i}r{\i}l{\i}nca yenisini almad{\i}lar. \\*
({\it The plate they broke was easily glued, so they didn't get a new one.})
\ex Yapt{\i}klar{\i} heykel ya\u{g}murda {\i}slan{\i}nca kurulamaya giri\c{s}tiler. \\*
({\it When the statue they made got wet in the rain, they started to dry it.})
\ex G{\i}d{\i}klad{\i}klar{\i} bebek \"{u}stlerine kusunca banyoya ko\c{s}tular. \\*
({\it When the baby they tickled vomited on them, they ran to the bathroom.})
\z
\ex {Ungrammatical Fillers}
\ea * K{\i}zd{\i}\u{g}{\i} bakan bulununca koltuk geri getirdi. \\*
({\it * When the minister whom he was angry at was found, the chair brought back.})
\ex * Arad{\i}\u{g}{\i} asistan ayr{\i}l{\i}nca proje birden unuttu. \\*
({\it * When the assistant he was looking for left, the project suddenly forgot.})
\ex * Bekledi\u{g}i kurye t\"{o}kezleyince soslar yere sa\c{c}t{\i}. \\*
({\it * When the courier he was waiting for stumbled, sauces spilled on the floor.})
\ex * G\"{u}ld\"{u}\u{g}\"{u} soytar{\i} as{\i}l{\i}nca \c{s}apka yerinde buldu. \\*
({\it * When the jester he laughed at was hanged, the hat found in its place.})
\ex * Yaz{\i}\c{s}t{\i}\u{g}{\i} dekan hap\c{s}urunca \c{c}aylar aniden d\"{u}\c{s}\"{u}rd\"{u}. \\*
({\it * The dean he corresponded with sneezed, the tea suddenly dropped.})
\ex * Bildi\u{g}i g\"{o}zl\"{u}k\c{c}\"{u} evlenince hediyeler a\u{g}lanarak verdi. \\*
({\it * When the optician he knew got married, gifts gave while crying.})
\ex * Savundu\u{g}u politikac{\i} yakalan{\i}nca a\c{c}{\i}klama haliyle kesti. \\*
({\it * When the politician he was defending was caught, the statement cut.})
\ex * A\u{g}{\i}rlad{\i}\u{g}{\i} temizlik\c{c}i bay{\i}l{\i}nca deterjan tekrar sa\c{c}t{\i}. \\*
({\it * When the housekeeper he was hosting fainted, the detergent spilled again.})
\ex * D\"{u}\c{s}ledi\u{g}i manken ni\c{s}anlan{\i}nca haber h{\i}zl{\i}ca yayd{\i}. \\*
({\it * The news spread quickly when the mannequin of her dreams got engaged.})
\ex * S\"{o}zle\c{s}ti\u{g}i fabrikat\"{o}r vurulunca kamera sessizce s\"{o}kt\"{u}. \\*
({\it * The camera silently dismantled when the fabricator he engaged with was shot.})
\ex * {\.I}stedi\u{g}i dans\"{o}z var{\i}nca kap{\i} sakince a\c{c}t{\i}. \\*
({\it * The door calmly opened when the belly dancer he wanted arrived.})
\ex * Haberle\c{s}ti\u{g}i \c{c}evirmen aramay{\i}nca metin keyfince bitirdi. \\*
({\it * When the translator with whom he communicated did not call, the text ended arbitrarily.})
\ex * Ba\u{g}{\i}rd{\i}\u{g}{\i} muhasebeci kovulunca hesap tamamen kar{\i}\c{s}t{\i}rd{\i}. \\*
({\it * The account completely messed up when the accountant he shouted was fired.})
\ex * Bulu\c{s}tu\u{g}u k\"{u}rk\c{c}\"{u} d\"{o}n\"{u}nce kuma\c{s} erkenden dikti. \\*
({\it * When the furrier he met returned, the fabric sewed early.})
\ex * Se\c{c}ti\u{g}i rekt\"{o}r atan{\i}nca k\"{u}t\"{u}phane gece kapatt{\i}. \\*
({\it * When the rector he chose was appointed, the library closed at night.})
\ex * G\"{o}r\"{u}\c{s}t\"{u}\u{g}\"{u} \c{s}of\"{o}r gecikince trafik aniden kilitledi. \\*
({\it * When the driver he was seeing was late, the traffic suddenly blocked.})
\ex * Kand{\i}rd{\i}\u{g}{\i} adam \"{o}demeyince bula\c{s}{\i}k saatlerce y{\i}kad{\i}. \\*
({\it * When the man he deceived did not pay, the dishes washed for hours.})
\ex * Mesajla\c{s}t{\i}\u{g}{\i} tesisat\c{c}{\i} gelince borular g\"{u}\c{c}l\"{u}kle s\"{o}kt\"{u}. \\*
({\it * When the plumber he was texting came in, the pipes unscrewed with difficulty.})
\ex * \"{U}zd\"{u}\u{g}\"{u} mobilyac{\i} k{\i}z{\i}nca koltuk sinirle par\c{c}alad{\i}. \\*
({\it * When the furniture maker he upset got angry, the sofa smashed angrily.})
\ex * Tan{\i}\c{s}t{\i}\u{g}{\i} medyum yan{\i}l{\i}nca fincan \"{o}fkeyle k{\i}rd{\i}. \\*
({\it * When the psychic he met was wrong, the cup broke angrily.})
\z
\z

\newpage