


\chapter{EXPERIMENT 4 ITEMS} \label{ap:exp4items}

\section{Experimental items}

The experimental items used in this experiment were the same as those in Experiment 2A except for a post-verbal sentence final additional of either \emph{efendim} (sir) or \emph{lan} (dude) depending on the condition. 

\section{Sub-Experiment 1 items}

\ea E\u{g}er zaman{\i}nda defteri veya kitab{\i} okursam fark{\i} anlar{\i}m. \\*
({\it I can tell the difference if I read the notebook or book in time.})
\ex E\u{g}er mutfakta elmay{\i} veya armudu y{\i}karsam g\"{u}zelce kurular{\i}m. \\*
({\it If I wash the apple or pear in the kitchen, I dry them well.})
\ex E\u{g}er una vanilyay{\i} veya tuzu eklersem iyice kar{\i}\c{s}t{\i}r{\i}r{\i}m. \\*
({\it If I add vanilla or salt to the flour, I mix it well.})
\ex E\u{g}er geceleyin Ven\"{u}s\"{u} veya Mars{\i} g\"{o}r\"{u}rsem foto\u{g}raf \c{c}ekerim. \\*
({\it If I see Venus or Mars at night, I'll take a picture.})
\ex E\u{g}er ma\u{g}azada pantolonu veya g\"{o}mle\u{g}i be\u{g}enirsem c\"{u}zdan{\i}m{\i} yoklar{\i}m. \\*
({\it If I like the trousers or the shirt in the store, I will check my wallet.})
\ex E\u{g}er pikni\u{g}e mangal{\i} veya k\"{o}m\"{u}r\"{u} getirirsem herkesi sevindiririm. \\*
({\it If I bring a barbecue or charcoal to the picnic, I'll make everyone happy.})
\ex E\u{g}er \c{s}irkette yaz{\i}c{\i}y{\i} veya taray{\i}c{\i}y{\i} kullan{\i}rsam ka\u{g}{\i}tlar{\i} toplar{\i}m. \\*
({\it If I use the printer or scanner at the company I collect the papers.})
\ex E\u{g}er sofraya pilav{\i} veya fasulyeyi koyarsam herkese payla\c{s}t{\i}r{\i}r{\i}m. \\*
({\it If I put rice or beans on the table, I'll divide it for everyone.})
\ex E\u{g}er g\"{o}revlilere yang{\i}n{\i} veya depremi bildirirsem adresimi hat{\i}rlat{\i}r{\i}m. \\*
({\it If I report a fire or an earthquake to the officials, I will remind them of my address.})
\ex E\u{g}er hastaya merhemi veya bandaj{\i} uygularsam ate\c{s}ini \"{o}l\c{c}erim. \\*
({\it If I apply the ointment or bandage to the patient, I take his temperature.})
\ex E\u{g}er doktoru hastaneye veya bakanl{\i}\u{g}a \c{s}ikayet edersem dilek\c{c}e yazar{\i}m. \\*
({\it If I report the doctor to the hospital or ministry, I write a petition.})
\ex E\u{g}er tatilde konaklamaya veya yeme\u{g}e kar{\i}\c{s}{\i}rsam bana k{\i}zarlar. \\*
({\it They get mad at me if I mess with accommodation or food on vacation.})
\ex E\u{g}er okulda \"{o}\u{g}retmene veya \"{o}\u{g}renciye sinirlenirsem duda\u{g}{\i}m{\i} {\i}s{\i}r{\i}r{\i}m. \\*
({\it If I get angry at a teacher or student at school, I bite my lip.})
\ex E\u{g}er yaz{\i}n k\"{o}ye veya tatile gidebilirsem \c{c}ok e\u{g}leniyorum. \\*
({\it I have a lot of fun if I can go to the village or on vacation in the summer.})
\ex E\u{g}er problemi asistana veya profes\"{o}re sorarsam cevab{\i} al{\i}r{\i}m. \\*
({\it If I ask the problem to the assistant or professor, I get the answer.})
\ex E\u{g}er sar{\i}msa\u{g}{\i} salataya veya sosa atarsam iyice ezerim. \\*
({\it If I put the garlic in a salad or dressing, I mash it well.})
\ex E\u{g}er aynay{\i} yukar{\i}ya veya sa\u{g}a kald{\i}r{\i}rsam arkay{\i} g\"{o}r\"{u}yorum. \\*
({\it If I lift the mirror up or to the right, I see the back.})
\ex E\u{g}er masay{\i} duvara veya dolaba hizalarsam temizlemem laz{\i}m. \\*
({\it If I align the table against the wall or closet, I have to clean it.})
\ex E\u{g}er tercihi topluma veya ba\c{s}kas{\i}na b{\i}rak{\i}rsam gelece\u{g}imi planlayamam. \\*
({\it I can't plan my future if I leave the choice to the community or someone else.})
\ex E\u{g}er ko\c{s}ullar{\i} b\"{o}lgeye veya insana uyarlarsam uyum sa\u{g}lanacak. \\*
({\it If I adapt the conditions to the region or the person, the adaptation will be achieved.})
\ex E\u{g}er et dolapta veya buzlukta beklerse dinlenmi\c{s} olur. \\*
({\it If the meat stays in the fridge or freezer, it is rested.})
\ex E\u{g}er s\"{u}rg\"{u}ler \c{c}ekmecede veya rafta kullan{\i}l{\i}rsa fiyat artar. \\*
({\it If the sliders are used in a drawer or on a shelf, the price increases.})
\ex E\u{g}er m\"{u}zik arabada veya konserde dinlenirse daha e\u{g}lenceli. \\*
({\it It's more fun if the music is listened to in the car or at a concert.})
\ex E\u{g}er sinekler evde veya binada yuvalan{\i}rsa \c{c}o\u{g}almaya ba\c{s}larlar. \\*
({\it If flies nest in a house or building, they begin to multiply.})
\ex E\u{g}er yemek sokakta veya lokantada yenirse daha tatl{\i}. \\*
({\it The food is more delicious if eaten on the street or in a restaurant.})
\ex E\u{g}er sporcu antrenmanda veya ma\c{c}ta ko\c{s}arsa kondisyonu artar. \\*
({\it If the athlete runs in training or a match, his fitness improves.})
\ex E\u{g}er birikim d\"{o}vizde veya alt{\i}nda tutulursa kazan\c{c} sa\u{g}lan{\i}r. \\*
({\it If the savings are kept in foreign currency or below, a profit is made.})
\ex E\u{g}er zamlar elektrikte veya suda yo\u{g}unla\c{s}{\i}rsa \"{o}demeler gecikecek. \\*
({\it Payments will be delayed if the hikes are concentrated on electricity or water.})
\ex E\u{g}er bakteriler suda veya karanl{\i}kta beklerse yap{\i}lar{\i} de\u{g}i\c{s}ir. \\*
({\it If bacteria stay in water or in the dark, their structure changes.})
\ex E\u{g}er lekeler kulakta veya bo\u{g}azda yayg{\i}nla\c{s}{\i}rsa doktorunuza ba\c{s}vurun. \\*
({\it If the spots become widespread in the ears or throat, contact your doctor.})
\ex E\u{g}er bebek k\"{o}pekten veya kediden korkarsa kuca\u{g}{\i}na al. \\*
({\it If the baby is afraid of a dog or cat, hold it in your lap.})
\ex E\u{g}er hediye aileden veya arkada\c{s} gelirse \c{c}ok sevinirsin. \\*
({\it If the gift comes from family or friends, you will be very happy.})
\ex E\u{g}er ceza maa\c{s}tan veya bonustan kesilirse gelirin azalacak. \\*
({\it If the penalty is deducted from the salary or bonus, your income will decrease.})
\ex E\u{g}er ziyaret\c{c}i kap{\i}dan veya salondan ge\c{c}erse bizimle kar\c{s}{\i}la\c{s}acak. \\*
({\it If the visitor passes through the door or the hall, he will meet us.})
\ex E\u{g}er yazar bask{\i}dan veya sald{\i}r{\i}dan usan{\i}rsa \"{u}lkeyi terk eder. \\*
({\it If the author gets tired of the pressure or the attack, he leaves the country.})
\ex E\u{g}er \"{o}\u{g}renci internetten veya \"{o}\u{g}retmenden faydalan{\i}rsa geli\c{s}imini h{\i}zland{\i}r{\i}r. \\*
({\it If the student benefits from the internet or the teacher, it accelerates his/her development.})
\ex E\u{g}er \c{c}ocuklar televizyondan veya oyundan b{\i}karsa onlar{\i} gezdirebilirsin. \\*
({\it If the kids get tired of the TV or the game, you can give them a ride.})
\ex E\u{g}er ya\c{s}l{\i}lar s{\i}caktan veya nemden hay{\i}flan{\i}rsa klimay{\i} a\c{c}abilirsin. \\*
({\it If the elderly lament the heat or humidity, you can turn on the air conditioner.})
\ex E\u{g}er yanaklar{\i} so\u{g}uktan veya utanga\c{c}l{\i}ktan k{\i}zar{\i}rsa y\"{u}z\"{u}n\"{u} saklar. \\*
({\it If her cheeks turn red from cold or shyness, she hides her face.})
\ex E\u{g}er pantolon ketenden veya pamuktan yap{\i}l{\i}rsa alerji yapmayabilir. \\*
({\it If the trousers are made of linen or cotton, they may not cause allergies.})
\z

\section{Sub-Experiment 2 items}

\ea Duydu\u{g}uma g\"{o}re bana/ben ve/ya da Deniz'e mektup gelmi\c{s}. \\*
({\it According to what I heard, a letter came to me\textsubscript{DAT/NOM} and/or Deniz.})
\ex Falc{\i}ya g\"{o}re bana/ben ve/ya da Ekin'e yol \c{c}{\i}km{\i}\c{s}. \\*
({\it According to the fortune teller, there was a way for me\textsubscript{DAT/NOM} and/or Ekin.})
\ex Baksana Twitter'da bana/ben ve/ya da Olgun'a laf sokmu\c{s}. \\*
({\it Look, he made fun of me\textsubscript{DAT/NOM} and/or Olgun on Twitter.})
\ex Onlara de\u{g}il, bana/ben ve/ya da Lale'ye fatura kesilecek. \\*
({\it Not them, me\textsubscript{DAT/NOM} and/or Lale will be billed.})
\ex Bilmiyorum valla, bana/ben ve/ya da Galip'e kahve {\i}smarlayacakm{\i}\c{s}. \\*
({\it I don't know, he was going to buy me\textsubscript{DAT/NOM} and/or Galip coffee.})
\ex Disiplin kurulu bana/ben ve/ya da Elif'e ceza verecek. \\*
({\it Disciplinary committee will punish me\textsubscript{DAT/NOM} and/or Elif.})
\ex Haftan{\i}n sonunda bana/ben ve/ya da Metin'e bilet verecekler. \\*
({\it At the end of the week they will give me\textsubscript{DAT/NOM} and/or Metin a ticket.})
\ex Maile bak{\i}l{\i}rsa bana/ben ve/ya da \"{O}mer'e haddimizi bildirecek. \\*
({\it According to the mail, he will tell me\textsubscript{DAT/NOM} and/or Ömer our place.})
\ex Ma\c{c}tan sonra bana/ben ve/ya da \"{U}mit'e g\"{u}n\"{u}m\"{u}z\"{u} g\"{o}sterecekmi\c{s}. \\*
({\it After the match, he was going to show me\textsubscript{DAT/NOM} and/or Ümit our day.})
\ex Dedi\u{g}ine g\"{o}re bana/ben ve/ya da Furkan'a izin verecek. \\*
({\it It says it will give me\textsubscript{DAT/NOM} and/or Furkan a day off.})
\ex Yazd{\i}klar{\i}na bak{\i}l{\i}rsa bana/ben ve/ya da Muhammed'e s{\i}r{\i}ks{\i}klam a\c{s}{\i}km{\i}\c{s}. \\*
({\it According to what he wrote, he was deeply in love with me\textsubscript{DAT/NOM} and/or Mohammed.})
\ex \c{S}una baksana bana/ben ve/ya da Onur'a delicesine hayranm{\i}\c{s}. \\*
({\it Look at this, he's mad stan of me\textsubscript{DAT/NOM} and/or Onur.})
\ex Hep derdi bana/ben ve/ya da Beyzanur'a araba alacakm{\i}\c{s}. \\*
({\it He always said he would buy a car for me\textsubscript{DAT/NOM} and/or Beyzanur.})
\ex \"{O}n\"{u}m\"{u}zdeki g\"{u}nlerde bana/ben ve/ya da G\"{u}ne\c{s}'e engel olacakm{\i}\c{s}. \\*
({\it He will interfere with me\textsubscript{DAT/NOM} and/or G\"{u}ne\c{s} in the coming days.})
\ex Son zamanlarda bana/ben ve/ya da U\u{g}ur'a benzemeye \c{c}al{\i}\c{s}{\i}yor. \\*
({\it She's been trying to look like me\textsubscript{DAT/NOM} and/or Uğur lately.})
\ex Annem haftasonu bana/ben ve/ya da Ay\c{s}eg\"{u}l'e temizlik yapt{\i}rtacak. \\*
({\it My mom is going to have me\textsubscript{DAT/NOM} and/or Ayşegül clean up at the weekend.})
\ex Bunu \"{o}\u{g}rendiklerinde bana/ben ve/ya da Senem'e hesap soracaklar. \\*
({\it When they find out about this, they will call me\textsubscript{DAT/NOM} and/or Senem accountable.})
\ex Yapt{\i}klar{\i}m{\i}z{\i} g\"{o}r\"{u}nce bana/ben ve/ya da Taner'e d\"{u}\c{s}man kesilecek. \\*
({\it When they see what we have done, they will turn against me\textsubscript{DAT/NOM} and/or Taner.})
\ex D\"{u}\c{s}\"{u}nsene mesela bana/ben ve/ya da Firdevs'e davet g\"{o}nderiyormu\c{s}. \\*
({\it Think about it, he was sending an invitation to me\textsubscript{DAT/NOM} and/or Firdevs.})
\ex R\"{u}ylar{\i}mda hep bana/ben ve/ya da G\"{o}ktu\u{g}'a mesaj at{\i}yor. \\*
({\it He always texts me\textsubscript{DAT/NOM} and/or Göktuğ in my dreams.})
\ex Eninde sonunda bana/ben ve/ya da Y{\i}ld{\i}z'a yol g\"{o}r\"{u}necek. \\*
({\it Eventually, the path will appear to me\textsubscript{DAT/NOM} and/or Y{\i}ld{\i}z.})
\ex \"{O}nemli olan bana/ben ve/ya da Berkay'a zarar gelmemesi. \\*
({\it The important thing is that no harm comes to me\textsubscript{DAT/NOM} and/or Berkay.})
\ex Eve gelince bana/ben ve/ya da Serkan'a dert yanacakm{\i}\c{s}. \\*
({\it When he comes home, he will have trouble with me\textsubscript{DAT/NOM} and/or Serkan.})
\ex Baksana heralde bana/ben ve/ya da Ay\c{s}enur'a nutuk \c{c}ekecek. \\*
({\it I guess he's going to lecture me\textsubscript{DAT/NOM} and/or Ayşenur.})
\ex Heralde \c{c}ocu\u{g}unu bana/ben ve/ya da Mehmet'e \c{s}ikayet edecek. \\*
({\it She will probably complain about her child to me\textsubscript{DAT/NOM} and/or Mehmet.})
\ex {\.I}ngiltere'den getirdiklerini bana/ben ve/ya da G\"{o}khan'a hediye edecek. \\*
({\it He will present to me\textsubscript{DAT/NOM} and/or Gökhan what he brought from England.})
\ex Bilge dedemiz bana/ben ve/ya da Selin'e yol g\"{o}sterecek. \\*
({\it Our wise grandfather will guide me\textsubscript{DAT/NOM} and/or Selin.})
\ex Teyzem kuzenimi bana/ben ve/ya da Nesrin'e emanet edecek. \\*
({\it My aunt will entrust my cousin to me\textsubscript{DAT/NOM} and/or Nesrin.})
\ex Eskisi gibi bana/ben ve/ya da G\"{u}lay'a ilgi g\"{o}stermiyor. \\*
({\it He doesn't show interest in me\textsubscript{DAT/NOM} and/or G\"{u}lay like he used to.})
\ex Ofise gelince bana/ben ve/ya da G\"{u}ls\"{u}m'e emir verecekmi\c{s}. \\*
({\it When he comes to the office, he will give orders to me\textsubscript{DAT/NOM} and/or G\"{u}ls\"{u}m.})
\ex Bahsetti\u{g}i kitab{\i} bana/ben ve/ya da I\c{s}{\i}l'a kargoyla g\"{o}nderecekmi\c{s}. \\*
({\it He was going to send the book he mentioned to me\textsubscript{DAT/NOM} and/or I\c{s}{\i}l by courier.})
\ex Yar{\i}n ak\c{s}am bana/ben ve/ya da Yaren'e pasta yapacakm{\i}\c{s}. \\*
({\it He is going to bake a cake for me\textsubscript{DAT/NOM} and/or Yaren tomorrow night.})
\ex Bug\"{u}n yar{\i}n bana/ben ve/ya da Buket'e haber verebilirler. \\*
({\it They can inform me\textsubscript{DAT/NOM} and/or Buket today or tomorrow.})
\ex Olanlar{\i} \"{o}\u{g}renince bana/ben ve/ya da Bet\"{u}l'e gizlice s\"{o}yleyecekmi\c{s}. \\*
({\it He was going to tell me\textsubscript{DAT/NOM} and/or Bet\"{u}l secretly when he found out what happened.})
\ex Konferanstan \"{o}nce bana/ben ve/ya da Fatih'e sunum yapacakm{\i}\c{s}. \\*
({\it He was going to make a presentation to me\textsubscript{DAT/NOM} and/or Fatih before the conference.})
\ex Yak{\i}n zamanda bana/ben ve/ya da Yasin'e k{\i}smet \c{c}{\i}kacakm{\i}\c{s}. \\*
({\it Soon, There will be appropriate partners for me\textsubscript{DAT/NOM} and/or Yasin.})
\ex Maa\c{s}{\i}n{\i} al{\i}nca bana/ben ve/ya da Ecem'e bilgisayar alacakm{\i}\c{s}. \\*
({\it He was going to buy a computer for me\textsubscript{DAT/NOM} and/or Ecem when he got his salary.})
\ex Yazd{\i}\u{g}{\i} \c{s}ark{\i}y{\i} bana/ben ve/ya da Nur'a arma\u{g}an etmi\c{s}. \\*
({\it He dedicated the song he wrote to me\textsubscript{DAT/NOM} and/or Nur.})
\ex Patron haftaya bana/ben ve/ya da Eray'a dosyalar{\i} d\"{u}zenlettirecek. \\*
({\it The boss will have me\textsubscript{DAT/NOM} and/or Eray edit the files next week.})
\ex M\"{u}d\"{u}r galiba bana/ben ve/ya da Orkun'a i\c{s} kitleyecek. \\*
({\it I think the manager dump some job on me\textsubscript{DAT/NOM} and/or Orkun.})
\z
